A Transdiciplinary approach to developing a spectral numerical coastal-flow

### 2019-08-30 literature_review.Rmd
adding to measuring waves section

Waves are often thought of as an elevation of the sea surface from a specific point over a period of time but this is obvisiously not he case. This is known as *surface elevation* and is the instaneous elevation of the sea surface abouve a specific point in a time record (see figure \ref{fig:surface elevation figures}). Although surface elevation does not represent a wave, it can be used to create a wave profile. This is achieved by profiling the the surface elevation between two successive *downward zero-crossings* or *upwardward zero-crossings* (see figure \ref{fig:surface elevation figures}). Both zero-crossings are symmetrical and essentially the same statistically. However, in practice the downward zero-crossings are preferred as it takes steepness of a wave into account (the front, see figure \ref{fig:surface elevation figures}) which is relevant to characterising breaking waves. It should be noted that surface elevation can be negative while a wave profile cannot.

@@

Waves are complex and various approaches, techniques and devices have been developed in order to measure waves effectively. One of the ways that has been used extensivley in the past but less so today, are visual estimates which can be used to characterise *significant swell height* (Hs) and *significant swell period* (Tp). Visual estimates use 15 - 20 of the most well defined, higher waves of a number of wave groups to characterise Hs and Tp. Although these parameters are useful, they do not accurately reflect the waves that are occuring in nature. Ocean waves are a combination of wind sea (short, irregular, locally generated waves) and swell (long, smooth waves, generated by distant storms) and so more parameters are needed to seperate Hs and Tp driven by wind sea and swell, i.e. seperate Hs and Tp parameters for wind sea and swell. Even with seperate parameters for the different types of waves would still not be enough to effectivley characterise waves in the ocean. In order to characterise the detail and complexity of ocean waves a different approach must be used, know as the *spectral* approach. This approach is based on the idea that the sea surface can be characterised as the summation of a large number of harmonic wave components.

>>>>

Waves are often thought of as an elevation of the sea surface from a specific point over a period of time but this is obvisiously not he case. This is known as *surface elevation* and is the instaneous elevation of the sea surface abouve a specific point in a time record (see figure \ref{fig:surface elevation figures}). Although surface elevation does not represent a wave, it can be used to create a wave profile. This is achieved by profiling the the surface elevation between two successive *downward zero-crossings* or *upwardward zero-crossings* (see figure \ref{fig:surface elevation figures}). Both zero-crossings are symmetrical and essentially the same statistically. However, in practice the downward zero-crossings are preferred as it takes steepness of a approaching wave into account (the front, see figure \ref{fig:surface elevation figures}) which is relevant to characterising breaking waves. It is important to note that surface elevation can be negative while a wave profile cannot.

@@

Waves are complex and various approaches, techniques and devices have been developed in order to measure waves effectively. One of the ways that has been used extensivley used in the past but less so today, are visual estimates which can be used to characterise *significant swell height* (Hs) and *significant swell period* (Tp). Visual estimates use 15 - 20 of the most well defined, higher waves of a number of wave groups to characterise Hs and Tp. Although these parameters are useful, they do not accurately reflect the waves that are occuring in nature. Ocean waves are a combination of wind sea (short, irregular, locally generated waves) and swell (long, smooth waves, generated by distant storms) and so more parameters are needed to seperate Hs and Tp driven by wind sea and swell, i.e. seperate Hs and Tp parameters for wind sea and swell. However, even with seperate parameters for the different types of waves would still not be enough to effectivley characterise waves in the ocean. In order to characterise the detail and complexity of ocean waves a different approach must be used, know as the *spectral* approach. This approach is based on the idea that the sea surface can be characterised as the summation of a large number of harmonic wave components, known as the *random-phase/amplitude* model.

The wave spectrum is based on several spectra, namely the amplitude spectrum, variance spectrum, variance density spectrum (discontinous) and the variance density spectrum (continuous). A wave spectrum can be used to describe the sea surface as a stochastic process, however two conditions (assumptions) must first be met 1) the observation must be stationary and 2) the surface elevations must be Gaussian distributed. As mentioned previously, the random-phase/amplitude model approach characterises waves as harmonic and the time-record can be reproduced as the sum of a large number of harmonic waves using a *Fourier analysis* method. A Fourier analysis allows values of the amplitude and phase for each frequency to be characterised, which in turn provides the amplitude and phase spectrum for a single time-record. The amplitude spectrum is can be calculated by averaging many time-records which determines the *average amplitude spectrum*. This approach allows the entire wave record to be characterised but is not statistically relevant and therefore is not applicable to hydrodynamic modelling. In order for to attain a statistically relevant value, the variance of each wave component must be calculated. The variance is regarded as statistically relevant as the sum of the variances (i.e. random surface elevation) of the wave components is equal to the variance of the sum of the wave components. In addition, LWT dictates that the energy of waves is proportional to the variance and is therefore an indirect measure of wave energy. This indirect measurment can also be used to determine other important wave components such as wave-induced particle velocity and pressure variations. Important to note that the measurement of variance are discrete and is therefore limited in characterising waves. A continious scale is more applicable for hydrodynamic modelling purposes and is also closer to representing ocean waves and can be calculated by allowing the frequency interval () to approach zero.
